---
title: "Gov 50 Final Project"
author: "Stephany Gutierrez"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Introduction Section

In this study, I plan to examine the attitudes of Latinos towards immigration in the United States. My research question is do the factors of knowing someone is undocumented, being AfroLatino, and and worrying of a family member or close friend being deported result in a lower ranking in the question of establishing a way for most immigrants currently in the country illegally to stay in the United States legally? In this data set, a respondents attitude towards establishing a way for most immigrants currently in the country illegally to stay in the United States legally were ranked as 1 = Very important goal, 2 = Somewhat important goal, 3 = Not too important goal, or 4 = Not at all important goal, meaning a lower ranking is of higher importance and vice-versa. I hypothesize that the factors of knowing someone is undocumented or worrying of a family member or close friend being deported will result in a lower ranking of the goals importance. I also hypothesize being AfroLatino will also have a higher ranking than not being AfroLatino because predominantly black Latin America countries are not experiencing increases in immigration unlike countries from Central America. 

## Data Description

The data I will be using is from the Pew Research Center's American Trends Panel Wave 86, the National Survey of Latinos. The data was collected via self-administered web surveys. The surveys were conducted in both English and Spanish. The questionnaire was sent out through March 15th through March 28th 2021. and the research design is cross-sectional. The specific report I am using is based on those who self-identified as Hispanics in this web survey. Most of the questions were yes/ no or on a scale of 1-4. When a question was left unanswered or was skipped, a 99 was placed to identify this. There was no specific outcome of interest in this web survey as the pain purpose was just to document trends of American society. 

The data set can be found here:
"American Trends Panel Wave 86." Pew Research Center, Washington, D.C. (March 15 – March 28, 2021) https://www.pewresearch.org/hispanic/dataset/american-trends-panel-wave-86/.


I will be using the following variables to test my hypothesis: 


| Name:              | Description:                                                                         |
|--------------------|--------------------------------------------------------------------------------------|
| `QKEY`             | Unique ID number of person taking the web survey                                     |
| `KNOWUNAUTH_W86`   | Does the person know someone who is undocumented (1), or not (2)?                    |
| `AFROLATINO_W86`   | Does the person identity as Afro-Latinx or Afro-Caribbean (1), or not (2)?           |
| `IMMVAL_c_W86`     | Person's attitude towards establishing a way for most immigrants currently in the country illegally to stay in the United States legally  ( 1 = Very important goal, 2 = Somewhat important goal, 3 = Not too important goal, 4 = Not at all important goal)|
| `WORRYDPORT_W86` ` | Person's worry of a family member or close friend being deported ( 1 = A lot, 2 = Some, 3 = Not much, 4 = Not at all)                                                                                                        |




```{r}
library("haven")
library(tidyverse)
library(dplyr)
library(broom)

latinos <- read_sav("data/ATP W86.sav")

latinos <- latinos|>
  select(QKEY,KNOWUNAUTH_W86,AFROLATINO_W86, WORRYDPORT_W86, IMMVAL_c_W86)

latinos$KNOWUNAUTH_W86[latinos$KNOWUNAUTH_W86 == 99] <- NA
latinos$AFROLATINO_W86[latinos$AFROLATINO_W86 == 99] <- NA
latinos$IMMVAL_c_W86[latinos$IMMVAL_c_W86 == 99] <- NA
latinos$WORRYDPORT_W86[latinos$WORRYDPORT_W86 == 99] <- NA

latinos <- latinos |>
  na.omit(latinos)

```

## Plot that Summarizes the Dependent Variable:
```{r}
 latinos <- latinos |> mutate(
    attitudes = case_when(
      IMMVAL_c_W86 == 1 ~ "Very important goal",
      IMMVAL_c_W86 == 2 ~ "Somewhat important goal",
      IMMVAL_c_W86 == 3 ~ "Not too important goal",
      IMMVAL_c_W86 == 4 ~"Not at all important goal"
    )
    ) 

latinos_plot <- latinos |>
  ggplot(aes(x = attitudes)) +
  geom_bar() +
  labs(
    x = "Level of Importance",
    y = "Number of People who responded",
    title = "Latino's Attitudes Towards Establishing Pathway To Legal Immigration "
  )
latinos_plot

attitudes_table <- latinos |>
  group_by(attitudes) |>
  summarize(Count = n())
knitr::kable(attitudes_table)

```

1672 Latinos said they felt that establishing a way for most immigrants currently in the country illegally to stay in the United States legally was very important goal, 1,027 said it was a somewhat important goal and 350 said it was a not too important goal and 196 said it was not at all an important goal. Because there was 3,245 out of the 3375 Latinos surveyed answered this question that means that 51%, the majority of Latinos believe that establishing a way for most immigrants currently in the country illegally to stay in the United States legally was a very important goal.




## Results Section:

## AfroLatino Idenity Impact on Immigration Attitudes
```{r}
latinos <- latinos |>
  mutate(`AfroLatino` = if_else(AFROLATINO_W86 == 1, "AfroLatino", "Not AfroLatin"))

AfroLatino_table <- latinos |>
  group_by(AfroLatino, IMMVAL_c_W86) |>
  summarize(n = n()) |>
  mutate(prop = n /sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = AfroLatino, values_from = prop
  )
knitr::kable(AfroLatino_table, digits = 2)

afrolatino_means <- latinos |>
  group_by(AfroLatino) |>
  summarize(IMMVAL_c_W86_afro = mean(IMMVAL_c_W86))
afrolatino_means

library("reshape2")
data_means <- melt(afrolatino_means)                                     
head(data_means)  

plot_latinos <- ggplot(data = data_means, mapping = aes(x = AfroLatino, y = value)) +
geom_col(mapping = aes(fill = AfroLatino)) +
  labs(x = "AfroLatino Idenity",
y = "Average Survey Response to Immigration Question",
title = "Average Afro-Latino Survey Responses to Immigration Question")
plot_latinos
```

There is little to no impact between Latinos and AfroLatino in terms of attitudes towards immigration. They are almost identical, with the biggest difference being in the "Somewhat important goal" category with  AfroLatino selecting it at 34% vs Latinos selecting it with 31%.

## Knowing Someone Undocumented Impact's on Immigration Attitudes
```{r}
latinos <- latinos |>
  mutate(`Know_Undocumented` = if_else(KNOWUNAUTH_W86 == 1, "Knows Undocumented Immigrant", "Does not Know Undocumented Immigrant"))
       
Know_table <- latinos |>
  group_by(KNOWUNAUTH_W86, IMMVAL_c_W86) |>
  summarize(n = n()) |>
  mutate(prop = n /sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = KNOWUNAUTH_W86, values_from = prop
  )
knitr::kable(Know_table, digits = 2)

know_means <- latinos |>
  group_by(Know_Undocumented) |>
  summarize(IMMVAL_c_W86_know = mean(IMMVAL_c_W86))
know_means

know_m_means <- melt(know_means)                                     
head(know_m_means)  

plot_know <- ggplot(data = know_m_means, mapping = aes(x = Know_Undocumented, y = value, fill = Know_Undocumented)) +
geom_col() +
  labs(x = "Does Respondent Know an Undocumented Immigrant",
y = "Average Survey Response to Immigration Question") + theme(axis.text.x = element_text(size = 6))
plot_know
```
As can be seen the majority of those who know an undocumented immigrant believe that establishing a way for most immigrants currently in the country illegally to stay in the United States legally is a very important goal or a somewhat important goal. With 57% of Latinos who know an Undocumented Immigrant believing that  it is a very important goal. In contrast to only 45% of those who do not know an Undocumented Immigrant believing that it is a very important goal. Similar results can be seen in terms of attitude, with those who know an an undocumented immigrant thinking it is more of an important goal. Knowing an undocumented immigrant resulted in about .22 difference in means, meaning that those who knew someone that was undocumented on average gave a .22 lower score on the scale. In this case the lower the number on the scale, the higher ranked the goal is.

## Linear Regression

```{r}
lm_know <- lm(IMMVAL_c_W86 ~ Know_Undocumented, data = latinos)
lm_know

lm_know |>
broom::tidy() |>
select(term, estimate) |>
knitr::kable(digits = 2)
```

The intercept is 1.4 which means that for the variable of knowing/not knowing an undocumented immigrant the survey response is predicted to be 1.4 in between very important goal and somewhat important. here is a casual effect because the coefficient on knowing an undocumented immigrant the variable is -0.22, which means that when a Latino knows someone undocumented their survey response is predicted to be about lower on average. Thus, when a Latino knows someone that is undocumented they are more likely to survey that establishing a way for most immigrants currently in the country illegally to stay in the United States legally is more of an important goal. This because a lower score means that the priority of the goal is higher. T



## Worrying of Deportation Impact's on Immigration Attitudes
```{r}

latinos <- latinos |>
  mutate(
    worry = case_when(
      WORRYDPORT_W86 == 1 ~ "A lot",
      WORRYDPORT_W86 == 2 ~ "Some",
      WORRYDPORT_W86 == 3 ~ "Not much",
      WORRYDPORT_W86 == 4 ~"Not at all"
    )
    ) 
       
worry_table <- latinos |>
  group_by(WORRYDPORT_W86, IMMVAL_c_W86) |>
  summarize(n = n()) |>
  mutate(prop = n /sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = WORRYDPORT_W86, values_from = prop
  )
knitr::kable(worry_table, digits = 2)

worry_means <- latinos |>
  group_by(worry) |>
  summarize(IMMVAL_c_W86_worry = mean(IMMVAL_c_W86))
worry_means

worry_m_means <- melt(worry_means)                                     
head(worry_means)  

worry_plot <- ggplot(data = worry_m_means, mapping = aes(x = worry, y = value, fill = worry)) +
geom_col() +
  labs(x = "Does Respondent Worry About a Loved One Being Deported",
y = "Average Survey Response to Immigration Question") + theme(axis.text.x = element_text(size = 6)) + theme(axis.title = element_text(size = 10))
worry_plot
```
The bar plot shows that the more a Latino worries about a loved one being deported the lower they will score the immigration question meaning that they are more likely to rank the goal of  establishing a way for most immigrants currently in the country illegally to stay in the United States legally as more important. For those who worry a lot of a loved ones deportation, they are on average likely to rank the goal right below 1.25 meaning that it is a very important goal for them. For those who worry some of a loved ones deportation, they are on average likely to rank the goal right below 1.5 meaning that it is a between a very important or somewhat important goal for them. For those who worry not much of a loved ones deportation, they are on average likely to rank the goal right below 1.75 meaning they rank the goal lower than the other two groups. Those that not at all worry about a loved ones deportation are the group that is most likely to rank the goal the lowest in terms of priority.


## Final Results
```{r}

data <- read.table(text="
Variable, Avergae.Response.to.IMMVA.Q
AfroLatino, 1.691318
Not AfroLatino, 1.715743	
Knows Undocu, 1.616111
Does Not Know Undocu, 1.834022
Worries A lot, 1.223553
Worries Some, 1.481054
Worries Not Much, 1.702060
Worries Not at all, 1.973562", header=TRUE, sep=",")
data


all_data_plot <- ggplot(data, aes(x = Variable, y = Avergae.Response.to.IMMVA.Q, fill = Variable)) +
    geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(size = 4)) +
  labs(x = "Different Variables",
y = "Average Survey Response to Immigration Question",
title = " Average Survey Responses to Immigration Question based on Different Variables") + theme(plot.title = element_text(size = 9)) + theme(axis.title = element_text(size = 9))
all_data_plot


all_data_scatter <- ggplot(data, aes(x = Variable, y = Avergae.Response.to.IMMVA.Q)) +
  geom_point(aes(colour = Variable)) + theme(axis.text.x = element_text(size = 5)) +
   labs(x = "Different Variables",
y = "Average Survey Response to Immigration Question",
title = "Average Survey Responses to Immigration Question based on Different Variables") + theme(plot.title = element_text(size = 9)) + theme(axis.title = element_text(size = 9))

all_data_scatter 
```

As the bar plot and scatter plot show the attitudes of Latinos towards establishing a way for most immigrants currently in the country illegally to stay in the United States legally, based on the factors of knowing someone is undocumented, being AfroLatino, and worrying of a family member or close friend being deported. My hypothesis of knowing someone is undocumented or worrying of a family member or close friend being deported will result in a lower ranking of goal importance was proved true. This because those who knows an undocumented person ranked the goal lower than those who do not know an undocumented person and in this data set the lower the ranking the more important the goal is to the respondent. Similarly, those who worried more about a loved ones deportation also ranked the goal lower. On the other hand being AfroLatino vs not being AfroLatino had little to no impact on the respondents ranking proving my other hypothesis false.


## Conclusion

Using the Pew Research Center's American Trends Panel Wave 86, the National Survey of Latinos I was able to test my hypotheses. My first hypothesis was that the factors of knowing someone is undocumented, being AfroLatino, and worrying of a family member or close friend being deported would result in a lower ranking in the question of establishing a way for most immigrants currently in the country illegally to stay in the United States legally. The respondents would rank the goal on a scale of 1-4 with 1 meaning the respondent thought it was a very important goal, 2 meaning it was a somewhat important goal, 3 meaning it was a not too important goal, and 4 meaning it was not at all important goal. My second hypothesis was that being AfroLatino would result in  a higher ranking than not being AfroLatino because predominantly black Latin America countries are not experiencing increases in immigration unlike countries from Central America, therefore immigration reform was not as big of importance. I was able to test both hypotheses. My first hypothesis was failed to be rejected and my second was rejected. Knowing an undocumented immigrant did result in a lower ranking than not knowing one. Similarly, the more worried the respondent was of a loved one being deported the lower the ranking of the goal. In contrast, being AfroLatino vs not being AfroLatino did not result in any major difference.  One of the limitations of the data set was that it was a self conducted survey meaning that the survey could have been randomly filled out or not filled out honestly. In the future the study could be improved by asking the questions in person and including more identification factors like income level and gender. 

