---
title: "Gov 50 Final Project"
author: "Stephany Gutierrez"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---


## Project thoughts

I will be exploring Latino's attitudes towards immigration in the United States.

In this study, I plan to examine the extent to which people get high and with what drugs. I hypothesize that older people will most frequently use drugs like crack and cocaine while younger people will use drugs like marijuana. I also expect that the more highly addictive drugs will be used with more frequency and among younger people as addictions tend to start off with younger people. If I observe that older people use drugs at a higher frequency, this would provide support for my hypothesis. If, on the other hand, I observe that younger people use drugs at a higher frequency, this would provide evidence against my hypothesis.


| Name:              | Description:                                                                         |
|--------------------|--------------------------------------------------------------------------------------|
| `KNOWUNAUTH_W86`   | Does the person know someone who is undocumented (1), or not (2)?                    |
| `AFROLATINO_W86`   | Does the person identity as Afro-Latinx or Afro-Caribbean (1), or not (2)?           |
| `CHILD_LT12_W86`   | Is the person a parent or guardian of any children under the age 12 (1), or not (2)? |
| `ST2_W86`          | Person's attitude towards immigration ( 1 = Requires no change, 2 = Requires only minor changes, 3 = Requires major                        changes, 4 = Needs to be completely rebuilt)                                         |
| `WORRYDPORT_W86` ` | Person's worry of a family member or close friend being deported ( 1 = A lot, 2 = Some, 3 = Not much, 4 = Not at all)                                                                                                        |





```{r}
library("haven")
library(tidyverse)
library(dplyr)
library(broom)

ATP_W86 <- read_sav("data/ATP W86.sav")

attitude <- ATP_W86

latinos <- attitude|>
  select(QKEY,KNOWUNAUTH_W86,AFROLATINO_W86,CHILD_LT12_W86,ST2_W86, WORRYDPORT_W86)

latinos$KNOWUNAUTH_W86[latinos$KNOWUNAUTH_W86 == 99] <- NA
latinos$AFROLATINO_W86[latinos$AFROLATINO_W86 == 99] <- NA
latinos$CHILD_LT12_W86[latinos$CHILD_LT12_W86 == 99] <- NA
latinos$ST2_W86[latinos$ST2_W86 == 99] <- NA
latinos$WORRYDPORT_W86[latinos$WORRYDPORT_W86 == 99] <- NA

latinos <- latinos |>
  na.omit(latinos)


```


```{r}
latinos <- latinos |>
  mutate(AfroLatinx = if_else(AFROLATINO_W86 == 1, "AfroLatinx", "Not AfroLatinx"))

AfroLatinx_table <- latinos |>
  group_by(AfroLatinx) |>
  summarize(n = n()) |>
  mutate(prop = n /sum(n)) |>
  select(-n) |>
  pivot_wider(names_from = AfroLatinx, values_from = prop
  )
knitr::kable(AfroLatinx_table, digits = 2)

    

```

```{r}
  ##mutate(Know_Undocumented = if_else(KNOWUNAUTH_W86 == 1, "Knows Undocumented Immigrant", "Does not Know Undocumented Immigrant"),
        ## AfroLatinx = if_else(AFROLATINO_W86 == 1, "AfroLatinx", "Not AfroLatinx"),
        ## Parent = if_else(CHILD_LT12_W86 == 1, "Parent of Child Under 12", "Not a Parent of Child Under 12"))

```

